\typeout{IJCAI-11 Instructions for Authors}

% These are the instructions for authors for IJCAI-11.
% They are the same as the ones for IJCAI-07 with superficical wording
%   changes only.

\documentclass{article}
% The file ijcai11.sty is the style file for IJCAI-11 (same as ijcai07.sty).
\usepackage{ijcai11}

% Use the postscript times font!
\usepackage{times}
\setlength{\parindent}{0em}

% \usepackage[style=authoryear,sorting=ynt]{biblatex}
% \usepackage{bibtex}
% the following package is optional:
%\usepackage{latexsym} 

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.


\title{Machine Learning for Polygenic risk score}
% \thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
\author{Vu-Lam DANG\\ 
Universite Grenoble Alpes\\
Grenoble, France \\
vu-lam.dang@etu.univ-grenoble-alpes.fr \\
\\
Supervised by: Dr Michael G. B. Blum.} % Mosig student

\begin{document}




\maketitle

{% Mosig student
  {\hbox to0pt{\vbox{\baselineskip=10dd\hrule\hbox
to\hsize{\vrule\kern3pt\vbox{\kern3pt
\hbox{{\small I understand what plagiarism entails and I declare that this report }}
\hbox{{\small is my own, original work. }}
\hbox{{\small Vu-Lam DANG, 26/04/2019:}}
\kern3pt
}\hfil%\kern3pt
\vrule
}\hrule}
}}
}


\begin{abstract}
The advent of genomics prediction as a viable diagnostic tool for certain disease have 
become a reality in the last decade. One particular interesting metric is called 
Polygenic Risk Score, which summaries the genetic. For this internship project, we 
are interested in applying advancement in the field of machine learning to improve the predictive power of PRS.
\end{abstract}

\section{Introduction}

\subsection{Polygenic Risk Score}
Polygenic Risk Score (PRS) \cite{Dudbridge2013} is a single perimeter metric
constructed from the weighted sum of associated alleles within each subject.
Given a pair of traits $Y = (Y 1 , Y 2)'$ 0 expressed as a weighted sum of m genetic
effects and a bias indicate environmental and unaccounted genetic effects:

$$Y = \beta' G + \epsilon = [\sum_{i=1}^{m} \beta_{i1} G_1 + \epsilon_1, \sum_{i=1}^{m} \beta_{i2} G_2 + \epsilon_2]$$
where $\beta$ is a $m*2$ weight matrix, $G$ is $m$ length vector of genetic markers 
and $\epsilon$ is a pair of random error that's independent of G.

The Polygenic Risk Score is defined as:

$$\hat{S} = \sum_{i=1}^{m} \beta_{i1} G_1$$
Association between a trait and its composite score highly implies there exist a
genetic signal among the markers, and the evidence of genetic effect when there is
no obvious candidate can be obtained. Currently, PRS has been used
for association testing rather than predicting complex traits \cite{Dudbridge2013}.

However, \cite{Inouye1883} conducted a GWAS on 1.7 million genetic
variants using UK Biobank database on Coronary Artery Disease, and found
a strong association between PRS and hazard ration for CAD; demonstrated
the power of genomics risk prediction to stratify individuals, and highlights the
possibility for genomics screening early in life to support risk prediction and
preventive treatment.

\subsection{Calculating $\beta$}
The centrepiece of calculating polygenic score is to determine the weight matrix $\beta$.

Originally, $\beta$ was calculated using Linear Regression. To some degree, this method
provides good predictive power \cite{Dudbridge2013}. However, this method overlooks some 
important genetic effects, such as Linkage Disequilibrium. Furthermore, all SNPs are 
used in the final calculation while not all are useful, lead to some loss in performance.

Clumping and Thresholding (C+T) is a derivative from simple Linear Regression PRS. 
In this method, only SNPs with p-values that are under a threshold are selected, and 
related SNPs are clumped together as a single input variable. 
We implemented this model in R using a package called \textit{bigsnpr} \cite{bigsnpr} 
that provide various functions for genomics computation.

To specifically combat Linkage Disequilibrium, LDPred provide a Python package and 
standalone application to calculate PRS, with Linkage Disequilibrium modelled into the computation 
\cite{LDpred}.

\textit{lassosum} \cite{doi:10.1002/gepi.22050} is an R package that provides PRS using a 
Penalized Regression known as LASSO. 
This method is the state of the art, and so far provide the best AUC in all methods 
tested ($AUC = 0.73$).

\section{Methodology}

\subsection{C+T PRS}
\subsection{Lassosum}
\subsection{}


\subsection{Dataset}

The models are trained on a sample dataset of 8000 data points, each with more than 
$600000$ SNPs. Atest dataset is also provided. The same dataset is also used to verify 
the result of other methods (described in section 2).

These datasets are provided by Florian Prive.

The computation on this work is written in R and Python. For R we specifically use 2 
package bigstasr and bigsnpr,which provide methods to handle large dataset by using disk 
swap \cite{bigsnpr}.

\section{Conclusion}

Polygenic Score has been widely studied as a good tool to identify and early
diagnose high risked individuals. In order to make it a viable tool for clinical usage,
improvement has to be made to decrease false negative and reduce the computational
complexity of the model. By incorporating machine learning technique, we hope to
eventually develop PRS into a powerful and accurate metric and tool to assess
and identify clinical risk factor.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{main}

\end{document}

